The 18.821 Mathematics Project Lab Report [Proofs]

### 2013-09-28 particle_definitions.tex
Remove jon-proofs.

\begin{definition}
A n-dimensional path $\gamma(t): \mathrm{R} \to \mathrm{R}^n$ is a function which maps a time $t \in \R$, $0 \ge t \ge T_{f, \gamma}$, to a position $\bvec{X} \in \mathrm{R}^n$. 
\end{definition}

@@

>>>>

\begin{definition}
A n-dimensional path $\gamma(t): \mathrm{R} \to \mathrm{R}^n$ is a function which maps a time $t \in \R$, $t \in [0, T_{f, \gamma}]$, to a position $\bvec{X} \in \mathrm{R}^n$. 
\end{definition}

@@

% -----------------------------------------------------------------------------

\subsection{Particle Motion in Polar Coordinates}

The motion of the 2-dimensional particles in this paper will typically be described in the polar coordinate system, shown in figure \ref{fig:polar-param}.

\includefig{polar_param.eps}{A particle moving in 2-dimensional polar coordinate system centered at B}{polar-param}

\begin{lemma}
The time derivative of $\theta$ is given by
\[
\frac{d\beta}{dt} = \frac{a_c}{s}
\]
\end{lemma}

\begin{proof}

\includefig{theta_deriv.eps}{Tangential acceleration}{theta-deriv}

If we look at a point, $p$, subject to only a centripetal acceleration, $a_c$, the change in $\bvec{v}$ over an infinitesimal time, $dt$, is shown in the figure \ref{fig:theta-deriv} (the two vectors, $\bvec{v}$ and $\bvec{v}+\bvec{dv}$, are superimposed). From the definition of $a_c$

\[
\frac{dv}{dt} = a_c
\]

So...
\[
\frac{d\beta}{dt} = \frac{a_c}{s}
\]
\end{proof}

The proof of the lemma in the case where $a_t$ is also nonzero is very similar, since the component of $\bvec{dv}$ in the direction $\bhat{v}$ is negligeable compared to $v$.

Returning again to figure \ref{fig:polar-param}, the following equations can be derived

\begin{align}
  \frac{dr(t)}{dt}& = -s(t) \, \cos\paren{\theta(t)}\\
  \frac{d\phi(t)}{dt}& = s(t) \, \sin\paren{\theta(t)}\\
  \frac{d\theta(t)}{dt}& = \frac{d\phi(t)}{dt} - \frac{a_c(t)}{s(t)}\label{eq:theta-deriv}\\
  \frac{d}{dt} \frac{dr(t)}{dt}& = -\frac{ds(t)}{dt} \cos\paren{\theta(t)} + s(t) \sin\paren{\theta(t)} \frac{d\theta(t)}{dt}
\end{align}

\begin{lemma}
  For a particle, $p$, with bounded centripetal and tangential acceleration, then:

  (1) The functions $\phi(t)$ and $\theta(t)$ are continuous.

  (2) For two times $t_1$ and $t_2$, s.t. $t_1 < t_2$, $\theta(t_1) > a$ and $\theta(t_2) < a$, then $\theta(t_c) = a$ for some $t_c \in [t_1, t_2]$.
\end{lemma}

\begin{proof}
First off, it should be noted that a solution to \ref{eq:theta-deriv} exists because the rhs is Lipschitz continuous. The proof of (1) follows directly from the fact that the derivative of $\theta(t)$ exist and is bounded. (2) is just a restatement of the intermediate value theorem.
\end{proof}

\begin{lemma}
  For a particle, $p$, with nonzero speed, and no centripetal acceleration for $t \ge t_0$, then
  \[
    \begin{cases}
      \theta(t) \to \pi \quad \text{as} \quad t \to \infty \qquad &\text{if} \;\; \theta(t_0) > 0\\
      \theta(t) = 0 \quad \text{for} \quad t \ge t_0 \qquad &\text{if} \;\; \theta(t_0) = 0
    \end{cases}
  \]
\end{lemma}

\begin{proof}
The second case is the easiest to check,

\end{proof}

### 2013-09-28 path_theorems.tex
Remove jon-proofs.

@@

\includefig{polar_param.eps}{Particle parameters in polar coordinates. B is center of coordinate system, and p is the particle.}{polar-param}

\begin{lemma}
  For a particle, $p$,
  \begin{align*}
    \frac{dr(t)}{dt} = -s(t) \, \cos\paren{\phi(t) + \theta(t)}\\
    \frac{d\phi(t)}{dt} = -s(t) \, \sin\paren{\phi(t) + \theta(t)}
  \end{align*}
\end{lemma}

\begin{lemma}\label{theta-deriv}
  For a particle, $p$,
  \[
  \frac{d\theta(t)}{dt} = -\frac{a_c(t)}{s(t)}
  \]
\end{lemma}

\begin{lemma}
  For a particle, $p$, with bounded centripetal and tangential acceleration, than the functions $\phi(t)$ and $\theta(t)$ are continuous. Furthermore, for two times $t_1$ and $t_2$, s.t. $t_1 \le t_2$, $\phi(t_1) + \theta(t_1) > a$ and  
\end{lemma}

The proof follows directly from the fact that the derivatives of $\phi(t)$ and $\theta(t)$ exist: \ref{theta-deriv}).

\begin{lemma}
  For a particle, $p$, with nonzero speed, and no centripetal acceleration for $t \ge t_0$, then
  \[
    \begin{cases}
      \phi(t) \to \theta(t_0) \quad \text{as} \quad t \to \infty \qquad &\text{if} \;\; \theta(t_0) > 0\\
      \phi(t) = 0 \quad \text{for} \quad t \ge t_0 \qquad &\text{if} \;\; \theta(t_0) = 0
    \end{cases}
  \]
\end{lemma}

>>>>

\begin{lemma}
If $a_t=0$ and $\norm{a_c} \le a_{c,max}$, then the minimum radius of curvature of a point trajectory is given by
\begin{equation}
R_{min} = \frac{v^2}{a_{c,max}}
\end{equation}
\end{lemma}

\begin{theorem}
For a particle, $p$, traveling along an optimal path, $\hat{\gamma}(t)$ between two points $X_1$ and $X_2$, 
\end{theorem}

\begin{lemma}
Every optimal trajectory is constructed from line segments and circular sections of radius $R_{min}$. Every point along an optimal trajectory where $a_t=0$ has $R = \infty$ or $R = R_{min}$.
\end{lemma}

If we look at a position, p, on a trajectory. We can align a rectangular coordinate system with this point, such that $\hat{y} = \hat{v}$.

\begin{align}
a_c& = \ddot{r} - r \dot{\phi}^2\\
%
a_t& = \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}}\\
%
0& = \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}}\\
%
r^2 \dot{\phi}& = constant
\end{align}

\qed

@@
